---
title: "Comprehensive Bias Assessment"
params:
  periodtype: "years"
  num: 1
  date_ranges:
    - [2000, 2005]
  norm: "yes"
  max_spat_uncert: 10000
  type: "count"
  res: 10000
  prev: "Yes"
  metric: "Coefficient of variation"
  log: "FALSE"
  min_periods: 2
  output: "density"
  shape: NULL
  nSamps: 50
  n_breaks: 50
  env_var_column: "road_density"
  env_data: !r readRDS("aux_file.rds")
  data: !r readRDS("reformatted_data.rds")
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: flatly
---

```{r setup for testing, include=FALSE}
params <- list(
  periodtype = "years",
  num = 1,
  date_ranges = list(c(2000, 2005)),
  norm = "yes",
  max_spat_uncert = 10000,
  type = "count",
  res = 10000,
  prev = "Yes",
  metric = "Coefficient of variation",
  log = "FALSE",
  min_periods = 2,
  output = "density",
  shape = NULL,
  nSamps = 50,
  n_breaks = 50,
  env_var_column = "road_density",
  env_data = readRDS("aux_file.rds"),
  data = readRDS("reformatted_data.rds")
)

```

```{r loading libraries}
library(dplyr)
library(occAssess)
library(knitr)
library(ggplot2)
library(sf)
library(sp)
library(ggpubr)
library(raster)
library(patchwork)
```

## 1. Record Number Over Time

```{r clean-data}
cleaned_data <- params$data %>%
  filter(!is.na(year))
num_filtered <- nrow(params$data) - nrow(cleaned_data)
if (num_filtered > 0) {
  message(paste(num_filtered, "rows with NA values in the year column were removed."))
}
```

```{r define-periods}
if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    from <- range[1]
    to <- range[2]
    seq(from = from, to = to)
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}
```

```{r plot-record-number}
res <- assessRecordNumber(
  dat = cleaned_data,
  species = "species",
  periods = periods,
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = NULL,
  identifier = "identifier",
  normalize = ifelse(tolower(params$norm) == "yes", TRUE, FALSE)
)
res$plot
```

### Notes
The plot above shows the number of records per time period. Normalization adjusts counts to be comparable across different identifiers.

## 2. Species Number Over Time

```{r check-data}
if (!("spatial_uncertainty" %in% names(params$data))) {
  stop("This function requires the 'spatial_uncertainty' field in the dataset.")
}
```

```{r clean-data}
cleaned_data <- params$data %>%
  filter(!is.na(year))
num_filtered <- nrow(params$data) - nrow(cleaned_data)
if (num_filtered > 0) {
  message(paste(num_filtered, "rows with NA values in the year column were removed."))
}
```

```{r define-periods}
if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    from <- range[1]
    to <- range[2]
    seq(from = from, to = to)
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}
```

```{r plot-species-number}
res <- assessSpeciesNumber(
  dat = cleaned_data,
  species = "species",
  periods = periods,
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = "spatial_uncertainty",
  identifier = "identifier",
  maxSpatUncertainty = params$max_spat_uncert,
  normalize = ifelse(tolower(params$norm) == "yes", TRUE, FALSE)
)
res$plot
```

### Notes
The plot above shows the number of species per time period, normalized to help compare identifiers with different record counts.

## 3. Species Identification Over Time

```{r}
cleaned_data <- params$data %>%
  filter(!is.na(year))

if (!("spatial_uncertainty" %in% names(cleaned_data))) {
  stop("This function requires the 'spatial_uncertainty' field in the dataset.")
}

if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    seq(from = range[1], to = range[2])
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}

res <- assessSpeciesID(
  dat = cleaned_data,
  species = "species",
  periods = periods,
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = "spatial_uncertainty",
  identifier = "identifier",
  maxSpatUncertainty = params$max_spat_uncert,
  type = params$type
)

res$plot
```

## 4. Rarity Bias Over Time

```{r}
cleaned_data <- params$data %>%
  filter(!is.na(year))

if (!("spatial_uncertainty" %in% names(cleaned_data))) {
  stop("This function requires the 'spatial_uncertainty' field in the dataset.")
}

if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    seq(from = range[1], to = range[2])
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}

res_rarity <- assessRarityBias(
  dat = cleaned_data,
  species = "species",
  periods = periods,
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = "spatial_uncertainty",
  identifier = "identifier",
  maxSpatUncertainty = params$max_spat_uncert,
  res = params$res,
  prevPerPeriod = ifelse(params$prev == "Yes", TRUE, FALSE),
  metric = ifelse(params$metric == "Coefficient of variation", "r2", "cor")
)

res_rarity$plot
```

## 5. Spatial Coverage Over Time

```{r}
cleaned_data <- params$data %>%
  filter(!is.na(year))

if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    seq(from = range[1], to = range[2])
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}

spat_cov <- assessSpatialCov(
  dat = cleaned_data,
  periods = periods,
  res = params$res,
  logCount = as.logical(params$log),
  shp = params$shape,
  species = "species",
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = if (params$output != "density") "spatial_uncertainty" else NULL,
  maxSpatUncertainty = if (params$output != "density") params$max_spat_uncert else NULL,
  identifier = "identifier",
  output = ifelse(params$output == "density", "density", "nPeriods"),
  minPeriod = if (params$output != "density") params$min_periods else NULL
)

spat_cov <- lapply(spat_cov, function(p) {
  p + coord_equal() + theme_minimal() + theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
})

do.call(ggarrange, spat_cov)
```

## 6. Spatial Bias Over Time

```{r}
cleaned_data <- params$data %>%
  filter(!is.na(year))

if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) {
    seq(from = range[1], to = range[2])
  })
} else {
  periods <- sort(unique(cleaned_data$year))
}

mask <- rasterize(params$shape, raster(nrow = 1000, ncol = 1000, ext = extent(params$shape)))

res_bias <- assessSpatialBias(
  dat = cleaned_data,
  periods = periods,
  mask = mask,
  nSamps = params$nSamps,
  degrade = TRUE,
  species = "species",
  x = "x_coordinate",
  y = "y_coordinate",
  year = "year",
  spatialUncertainty = "spat_uncert",
  identifier = "identifier"
)

res_bias$plot
```

## 7. Environmental Bias Over Time

```{r}
convert_x_coordinate_y_coordinate_to_monad <- function(x_coordinate, y_coordinate) {
  grid_letters <- matrix(LETTERS[-9], ncol = 5, byrow = TRUE)[5:1, ]
  xo_500 <- trunc(x_coordinate / 500000) + 3
  yo_500 <- trunc(y_coordinate / 500000) + 2
  s1 <- grid_letters[cbind(yo_500, xo_500)]
  xo_100 <- trunc((x_coordinate %% 500000) / 100000) + 1
  yo_100 <- trunc((y_coordinate %% 500000) / 100000) + 1
  s2 <- grid_letters[cbind(yo_100, xo_100)]
  monad_e <- trunc((x_coordinate %% 100000) / 1000)
  monad_n <- trunc((y_coordinate %% 100000) / 1000)
  monads <- paste0(s1, s2, sprintf("%02d%02d", monad_e, monad_n))
  return(monads)
}

cleaned_data <- params$data %>%
  filter(!is.na(year))

if (params$periodtype == "ranges") {
  periods <- lapply(params$date_ranges, function(range) seq(from = range[1], to = range[2]))
} else {
  unique_years_range <- range(unique(cleaned_data$year), na.rm = TRUE)
  periods <- list(seq(from = unique_years_range[1], to = unique_years_range[2]))
}

positions <- cleaned_data %>%
  dplyr::select(x_coordinate, y_coordinate, year) %>%
  mutate(monad = convert_x_coordinate_y_coordinate_to_monad(x_coordinate, y_coordinate))

env_data <- params$env_data %>%
  filter(auxcolumn == params$env_var_column) %>%
  dplyr::select(monad, value) %>%
  rename(!!params$env_var_column := value)

for (i in seq_along(periods)) {
  monads_in_period <- positions %>%
    filter(year %in% periods[[i]]) %>%
    distinct(monad) %>%
    pull(monad)
  colname <- paste0("presence_", i)
  env_data[[colname]] <- ifelse(env_data$monad %in% monads_in_period, 1, 0)
}

plots <- lapply(seq_along(periods), function(i) {
  presence_col <- paste0("presence_", i)
  assessBias1D(
    pop = env_data,
    breaks = params$n_breaks,
    R = presence_col,
    x = params$env_var_column,
    RNames = presence_col
  )$plot
})

wrap_plots(plots, ncol = 1)
```

### Notes
These plots evaluate temporal and spatial bias metrics including sampling coverage, spatial clustering, and environmental representativeness.