---
title: "Spatial Coverage Report"
output: html_document
params:
  species: "species"
  periods: NULL
  x: "x_coordinate"
  y: "y_coordinate"
  year: "year"
  spatialUncertainty: NULL
  maxSpatUncertainty: 10000
  res: 1000
  logCount: FALSE
  identifier: "identifier"
  output: "density"
  minPeriod: 2
  shapefile_uploaded: FALSE
  country: NULL
  country_iso2: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(occAssess)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
```

## Input Parameters

- **Species**: `r params$species`
- **Periods**: `r paste(sapply(params$periods, function(p) paste(range(p), collapse = "-")), collapse = ", ")`
- **X Coordinate**: `r params$x`
- **Y Coordinate**: `r params$y`
- **Year**: `r params$year`
- **Max Spatial Uncertainty**: `r params$maxSpatUncertainty`
- **Resolution**: `r params$res`
- **Log Count**: `r params$logCount`
- **Identifier**: `r params$identifier`
- **Output**: `r params$output`
- **Minimum Periods**: `r params$minPeriod`
- **Shapefile Uploaded**: `r params$shapefile_uploaded`
- **Country**: `r params$country`

*Note: You can access the reformatted data by loading the `your_formatted_data.csv` file.*

```{r load-data}

data <- read.csv("export/your_formatted_data.csv")
```

```{r load-shape}
if (params$shapefile_uploaded) {
  shp_dir <- "export/user_shapefile_spatial_coverage"
  shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)
  if (length(shp_files) == 0) stop("Shapefile (.shp) not found in uploaded folder")
  shp_sf <- sf::st_read(shp_files[1])
  shp <- as(shp_sf, "Spatial")
} else {
  countriesLow <- readRDS("export/countriesLow.rds")
  shp <- countriesLow[countriesLow$ISO_A2 == params$country_iso2, ]
}
```

```{r assess-spatial-cov}
result <- assessSpatialCov(
  dat = data,
  periods = params$periods,
  res = params$res,
  logCount = params$logCount,
  shp = shp,
  species = params$species,
  x = params$x,
  y = params$y,
  year = params$year,
  spatialUncertainty = params$spatialUncertainty,
  maxSpatUncertainty = if (params$output == "density") NULL else params$maxSpatUncertainty,
  identifier = params$identifier,
  output = params$output,
  minPeriod = if (params$output %in% c("overlap", "nPeriods")) params$minPeriod else NULL
)

if (params$output == "density") {
  result <- lapply(result, function(p) {
    p + scale_fill_gradient(na.value = "white", low = "white", high = "blue") + coord_equal()
  })
} else if (params$output == "overlap") {
  result <- lapply(result, function(p) {
    p + coord_equal() + scale_fill_brewer(type = "qual") +
      theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
            axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
  })
} else {
  result <- lapply(result, function(p) {
    p + coord_equal() +
      theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
            axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
  })
}
```

```{r plot-results}
if (is.list(result)) {
  print(ggarrange(plotlist = result))
} else {
  print(result)
}
```
