---
title: "Species Rarity Bias Report"
output: html_document
params:
  species: "species"
  periods: NULL
  x: "x_coordinate"
  y: "y_coordinate"
  year: "year"
  spatialUncertainty: "spatial_uncertainty"
  identifier: "identifier"
  maxSpatUncertainty: 10000
  res: 10000
  prevPerPeriod: TRUE
  metric: "r2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(occAssess)
library(dplyr)
library(ggplot2)
```

## Input Parameters

- **Species**: `r params$species`
- **Periods**: `r paste(sapply(params$periods, function(p) paste(range(p), collapse = "-")), collapse = ", ")`
- **X Coordinate**: `r params$x`
- **Y Coordinate**: `r params$y`
- **Year**: `r params$year`
- **Spatial Uncertainty**: `r params$spatialUncertainty`
- **Identifier**: `r params$identifier`
- **Max Spatial Uncertainty**: `r params$maxSpatUncertainty`
- **Resolution**: `r params$res`
- **Prevalence Per Period**: `r params$prevPerPeriod`
- **Metric**: `r params$metric`

*Note: You can access the reformatted data by loading the `your_formatted_data.csv` file.*

```{r load-data}
data <- read.csv("export/your_formatted_data.csv")
```

```{r assess-rarity-bias}
result <- assessRarityBias(
  dat = data,
  species = params$species,
  periods = params$periods,
  x = params$x,
  y = params$y,
  year = params$year,
  spatialUncertainty = params$spatialUncertainty,
  identifier = params$identifier,
  maxSpatUncertainty = params$maxSpatUncertainty,
  res = params$res,
  prevPerPeriod = params$prevPerPeriod,
  metric = params$metric
)
```

```{r plot-results}
print(result$plot)
```
