---
title: "Species Bias Report"
output: html_document
params:
  species: "species"
  periods: NULL
  x: "x_coordinate"
  y: "y_coordinate"
  year: "year"
  spatialUncertainty: "spatial_uncertainty"
  identifier: "identifier"
  maxSpatUncertainty: 10000
  normalize: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}

library(dplyr)
library(ggplot2)

# install remotes to install occAssess
if(!require(remotes)){
  install.packages(remotes)
}

if(!require(occAssess)){
  install_github("robboyd/occAssess")
}

```

## Input Parameters

- **Species**: `r params$species`
- **Periods (Note: This is a list object)**: `r paste(sapply(params$periods, function(p) paste(range(p), collapse = "-")), collapse = ", ")`
- **X Coordinate**: `r params$x`
- **Y Coordinate**: `r params$y`
- **Year**: `r params$year`
- **Spatial Uncertainty**: `r params$spatialUncertainty`
- **Identifier**: `r params$identifier`
- **Max Spatial Uncertainty**: `r params$maxSpatUncertainty`
- **Normalize**: `r params$normalize`

```{r load-data}
data <- read.csv("export/your_formatted_data.csv")
```

```{r assess-species-number}
result <- assessSpeciesNumber(
  dat = data,
  species = params$species,
  periods = params$periods,
  x = params$x,
  y = params$y,
  year = params$year,
  spatialUncertainty = params$spatialUncertainty,
  identifier = params$identifier,
  maxSpatUncertainty = params$maxSpatUncertainty,
  normalize = params$normalize
)

# Save result
write.csv(result$data, file = "export/species_number_bias.csv", row.names = FALSE)

# Display plot
result$plot
```
