---
title: "Spatial Bias Report"
output: html_document
params:
  species: "species"
  periods: NULL
  x: "x_coordinate"
  y: "y_coordinate"
  year: "year"
  spatialUncertainty: "spat_uncert"
  identifier: "identifier"
  nSamps: 50
  shapefile_uploaded: FALSE
  country: NULL
  country_iso2: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}

library(dplyr)
library(raster)
library(sf)
library(sp)

# install remotes to install occAssess
if(!require(remotes)){
  install.packages(remotes)
}

if(!require(occAssess)){
  install_github("robboyd/occAssess")
}

```

## Input Parameters

- **Species**: `r params$species`
- **Periods (Note: This is a list object)**: `r paste(sapply(params$periods, function(p) paste(range(p), collapse = "-")), collapse = ", ")`
- **X Coordinate**: `r params$x`
- **Y Coordinate**: `r params$y`
- **Year**: `r params$year`
- **Spatial Uncertainty**: `r params$spatialUncertainty`
- **Identifier**: `r params$identifier`
- **Number of Simulations**: `r params$nSamps`
- **Shapefile Uploaded**: `r params$shapefile_uploaded`
- **Country**: `r params$country`

```{r load-data}
data <- read.csv("export/your_formatted_data.csv")
```

```{r load-shape}
if (params$shapefile_uploaded) {
  shp_dir <- "export/user_shapefile"
  shp_files <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE)
  if (length(shp_files) == 0) stop("Shapefile (.shp) not found in uploaded folder")
  shp_sf <- sf::st_read(shp_files[1])
  shp <- as(shp_sf, "Spatial")
} else {
  countriesLow <- readRDS("export/countriesLow.rds")
  shp <- countriesLow[countriesLow$ISO_A2 == params$country_iso2, ]
}
```

```{r raster-mask}
mask <- rasterize(shp, raster(nrow = 1000, ncol = 1000, ext = extent(shp)))
```

```{r assess-spatial-bias}
result <- assessSpatialBias(
  dat = data,
  periods = params$periods,
  mask = mask,
  nSamps = params$nSamps,
  degrade = TRUE,
  species = params$species,
  x = params$x,
  y = params$y,
  year = params$year,
  spatialUncertainty = params$spatialUncertainty,
  identifier = params$identifier
)

# Save the data
write.csv(result$data, file = paste0("export/spatial_bias.csv"), row.names = FALSE)

# View the plots
result$plot

```